\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-05-26 forwards-and-backwards.tex
Update on Overleaf.

\begin{lemma} \label{lem: tightness of individual spatial trajectory}
For fixed $a=(a_0,a_1,..,a_k)$,
$\{(X^{N}_{a}(t))_{t \geq \tau^{N}_a}: N \in \mathbb{N},  \}$ is relatively compact in 
$D_{\mathbb{R}^d}[0,\infty )$ where $\mathbb{R}^d$ is equipped with the standard Euclidean metric. 
\end{lemma}
\begin{proof}
First,
note that the birth time
of $a$ at the $N$-th scaling
regime, denoted as
$\tau^{N}_a$,
is identical in distribution
for all $N > 0$.
This is because the birth time
of $a=(a_0,a_1,...,a_k)$
comes after $a_1$ births
in the 1st generation,
followed by $a_2$ births
in the 2nd generation, etc..
Therefore $\tau^N_{a} \sim Gamma(\sum_{i=1}^{k}a_i,1)$.

Now we prove
convergence in distribution
of the birth locations $X^N_{a}(\tau^N_{a})$.
From Equation \eqref{eq: space_across_lineage},
we have
\begin{equation}
\begin{aligned}
&X^{N}_{[a]_{k}}(\tau^N_{a})- X^{N}_{[a]_{0}}(0)\\
=& \sum_{i=1}^{k}
\sum_{j=1}^{a_i-1}
(1-\kappa_{[a]_{i-1}\concat j}) y_{[a]_{i-1}\concat j} \ind_{U^N_{[a]_{i-1}\concat j}(\tau_{[a]_{i-1}\concat j}) < N}\\
&+ \sum_{i=1}^{k}
\kappa_{[a]_i} y_{[a]_i} \ind_{U^N_{[a]_i}(\tau_{[a]_i}) < N}.
\end{aligned}    
\end{equation}

Since $(s_{[a]_{i} \concat j}, v_{[a]_{i} \concat j}, y_{[a]_{i} \concat j}, \kappa_{[a]_{i} \concat j})
\in \Pi_{[a]_{i-1}}$
are all iid random variables 
for each $i$ and $j$,
and that the sum above is finite,
it suffices to show that
the quantities 
\begin{equation}
\mathbb{P}(U^N_{[a]_{i-1} \concat j}(\tau_{[a]_{i-1} \concat j}) < N)    
\end{equation}
converges
for all $i=1,..,k$ and $j=1,..,a_i$.

Writing out
$U^N_{[a]_{i-1} \concat j}(\tau_{[a]_{i-1} \concat j})$,
we have
\begin{equation}
\begin{aligned}
&\mathbb{P}(U^N_{[a]_{i-1} \concat j}(\tau_{[a]_{i-1} \concat j}) < N) \\
=&\mathbb{P}(U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j})
    +
    \ell(
    X^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j}-),
    y_{[a]_{i-1} \concat j},
    \eta^{N}(\tau_{[a]_{i-1} \concat j}-),
    v_{[a]_{i-1} \concat j})
    < N)\\
=&\mathbb{P}\left(
    \frac{
        v_{[a]_{i-1} \concat j}
    }{
        2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    }
    < N-U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j})
    \right)\\
=&\mathbb{P}\left(
            v^N_{[a]_{i-1} \concat j}
    < (2 \theta \gamma(x, \eta) r(x + y, \eta)-  2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j}))
    \right)\\
=& \frac{(2 \theta \gamma(x, \eta) r(x + y, \eta)-  2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j}))}{N}\\
\to & 2\alpha \gamma(x, \eta) r(x + y, \eta).
\end{aligned}    
\end{equation}

>>>>

\begin{lemma} \label{lem: tightness of individual spatial trajectory}
For fixed $a=(a_0,a_1,..,a_k)$,
$\{(X^{N}_{a}(t))_{t \geq \tau^{N}_a}: N \in \mathbb{N},  \}$ is relatively compact in 
$D_{\mathbb{R}^d}[0,\infty )$ where $\mathbb{R}^d$ is equipped with the standard Euclidean metric. 
\end{lemma}
\begin{proof}
First,
note that the birth time
of $a$ at the $N$-th scaling
regime, denoted as
$\tau^{N}_a$ converges to $0$
as $N \to \infty$.
This is because the birth time
of $a=(a_0,a_1,...,a_k)$
comes after $a_1$ births
in the 1st generation,
followed by $a_2$ births
in the 2nd generation, etc..
Furthermore,
these birth times are
independent of the environment. Finally,
we rescale time by a factor of $\theta$,
so $\tau^N_{a} \sim Gamma(\theta\sum_{i=1}^{k}a_i,1) \to 0$.

Now we prove
convergence in distribution
of the birth locations $X^N_{a}(\tau^N_{a})$.
From Equation \eqref{eq: space_across_lineage},
we have
\begin{equation}
\begin{aligned}
&X^{N}_{[a]_{k}}(\tau^N_{a})- X^{N}_{[a]_{0}}(0)\\
=& \sum_{i=1}^{k}
\sum_{j=1}^{a_i-1}
\ind_{\kappa_{[a]_{i-1}\concat j}=0}  \ind_{U^N_{[a]_{i-1}\concat j}(\tau_{[a]_{i-1}\concat j}) < N}y_{[a]_{i-1}\concat j}\\
&+ \sum_{i=1}^{k}
\ind_{\kappa_{[a]_{i-1}\concat j}=1} \ind_{U^N_{[a]_i}(\tau_{[a]_i}) < N}y_{[a]_i} .
\end{aligned}    
\end{equation}

Since $(s_{[a]_{i} \concat j}, v_{[a]_{i} \concat j}, y_{[a]_{i} \concat j}, \kappa_{[a]_{i} \concat j})
\in \Pi_{[a]_{i-1}}$
are all iid random variables 
for each $i$ and $j$,
we can give a simple bound
on $\mathbb{E}[(X^{N}_{[a]_{k}}(\tau^N_{a})- X^{N}_{[a]_{0}}(0))^2]$.

First of all, 
note that the total number
of jumps
the ancestral lineage 
has until time $\tau_{a}$
is at most $\sum_{i=1}^{k}a_i$.
Furthermore, each jump is independent and has mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
that is bounded uniformly across
the environment.

As a result, as $N \to \infty$,
\begin{equation}
\begin{aligned}
&\quad\mathbb{E}[(X^{N}_{[a]_{k}}(\tau^N_{a})- X^{N}_{[a]_{0}}(0))^2]\\
\leq & \quad\mathbb{E}\left[\left(\sum_{i=1}^{k}
\sum_{j=1}^{a_i-1} y_{[a]_{i-1}\concat j}+ \sum_{i=1}^{k}
y_{[a]_i}\right)^2\right]\\
\leq & \quad\mathbb{E}\left[\left(\sum_{i=1}^{k}
\sum_{j=1}^{a_i} y_{[a]_{i-1}\concat j}\right)^2\right]\\
=& \quad{ka_k \choose 2}\sup_{x \in \mathbb{R}^d}(||\meanq(x)||^2)/\theta^2 + ka_k \left(\sup_{x \in \mathbb{R}^d}(||\meanq(x)||^2)/\theta^2+\sup_{x \in \mathbb{R}^d}|tr(\covq(x) )|\right)\\
 \to &  \quad 0.
\end{aligned}    
\end{equation}

This is expected
as in the infinitesimal limit,
births are instantaneous. 
Therefore, all offspring 
within small number of generations
($k << \theta$)
from the original
ancestors are born
almost instantaneously
at the position of its
ancestor.
